$p$-adic precision

### 2014-01-30 padicprec.tex
Working on introduction to Section 2

>>>>

Two sources of inspiration arise when studying $p$-adic algorithms.
The first relates $\Zp$ to its quotients $\Z/p^n\Z$.  The preimage in
$\Zp$ of an element $a \in \Z / p^n \Z$ is a ball, and these balls cover
$\Zp$ for any fixed $n$.  Since the projection $\Zp \to \Z/p^n\Z$ is a
homomorphism, given unknown elements in two such balls we can
locate the balls in which their sum and product lie.  Working on a computer
we must find a way to write elements using only a finite amount of data.
By lumping elements together into these balls of radius $p^{-n}$, we may
model arithmetic in $\Zp$ using the finite ring $\Z / p^n \Z$.  In Sage
this method for approximating elements is known as ``fixed modulus''.

The second source draws upon parallels between $\Qp$ and $\R$.  Both
occur as completions of $\Q$, and both have dense subrings:
the localizations $\Z[\frac1p] \subset \Qp$ and $\Z[\frac12] \subset \R$.
In $\R$, floating point arithmetic provides approximate operations $\oplus$ and $\odot$ on a
subset $S \subset \Z[\frac12]$ that model $+$ and $\cdot$ in $\R$ up to a given relative precision $h$:
\[
\frac{x \circledast y}{x \ast y} - 1 \le 2^{-h}
\]
for all $x, y \in S$ and $\ast \in \{+, \cdot\}$.  Interval arithmetic in the reals
is unwieldy, since arithmetic operations always increase the lengths of the inputs.
As a consequence, most computations in the real numbers rely on
statistical cancelation and external estimates of precision loss, rather than
attempting to track known precision at each step.  This tendency is strengthened
by the ubiquity of floating point arithmetic in scientific applications, where Gaussian
distributions are more common than intervals anyway.

In the $p$-adic world, these features are reversed.
